---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

```{r, message=F, warning=F}
rm(list=ls())
set.seed(123)
library(tidyverse)
library(ggplot2)
library(DT)
library(VennDiagram)
library(rprojroot)
library(dplyr)
root <- rprojroot::find_rstudio_root_file()
```



## BrainSpan data analysis 


* [Run COBS on BrainSpan data](COBS_brainspan.html) 


## Input data 

1. TADA p values, FDRs for all genes in  Fu, et al paper and FDR's in Trost paper 

```{r, message=F, warning=F}
jack_fu=as_tibble(read.csv(file.path(root, "../../../Dataset/20231221_PMID_35982160/Fu 2022 Suppl-41588_2022_1104_MOESM3_ESM.csv")))  # 
jack_fu_185_asd_genes=jack_fu %>% filter(ASD185=="TRUE") %>% dplyr::select(gene_gencodeV33, gene_id)

length(jack_fu$gene_gencodeV33)
sum(jack_fu_185_asd_genes$gene_gencodeV33 %in% jack_fu$gene_gencodeV33)


trost_data=as_tibble(read.table("C:/han/Dataset/20250114_PMID_36368308_trost/TADA+_ASD_gene_list.all.tsv", header = TRUE, sep = "\t", stringsAsFactors = FALSE))

length(trost_data$gene)
sum(jack_fu_185_asd_genes$gene_gencodeV33 %in% trost_data$gene)


length(intersect(jack_fu$gene_gencodeV33, trost_data$gene))
```


* 18159 genes in fu data, and 17348 in trost data, 16143 are overlapping with each other 

* among 185 risk genes in fu data, 182 in trost data. 






2. Brainspan from Brueggeman, et al paper 


```{r, message=F, warning=F}
load("C:\\han\\Projects\\IntegratePPI\\forecASDPaper\\01_BrainSpan_matrix.RData") # bs is the brainspan data in foreASD paper
sfari_gene_ids=as_tibble(read.table("C:\\han\\Dataset\\20240610_PMID_33244169_forecASD\\sfari_gene_ids.txt", header=T, sep = "\t", fill = TRUE, stringsAsFactors = FALSE))
meta = read.csv("C:\\han\\Dataset\\20240610_PMID_33244169_forecASD\\composite_table.csv",stringsAsFactors = F, row.names = 1)  # tada data is in meta


#bs_genes=sapply(rownames(bs), function(x) (sfari_gene_ids$Gene.name[which(sfari_gene_ids$Gene.stable.ID==x)]))
bs_genes=unlist(sapply(rownames(bs), function(x) (meta$symbol[which(meta$ensembl_string==x)])))


length(intersect(bs_genes, jack_fu_185_asd_genes$gene_gencodeV33))
length(intersect(bs_genes, jack_fu$gene_gencodeV33))


```

* among 185 risk genes in fu paper, 178 appear in brainspan data 

* of 18159 genes in fu data, 15999 appear in brainspan data 



3. protein interaction from STRING, Brueggeman, et al paper


```{r, message=F, warning=F}
load("C:\\han\\Projects\\IntegratePPI\\forecASDPaper\\tmp01_STRING_graph.Rdata")

string_genes=unlist(sapply(rownames(string.path), function(x) (meta$symbol[which(meta$ensembl_string==x)])))


length(intersect(string_genes, jack_fu_185_asd_genes$gene_gencodeV33))
length(intersect(string_genes, jack_fu$gene_gencodeV33))

```

* 178/185 risk genes in fu paper, with a total of 15999 overlapping genes.   



4. forecASD results 

```{r, message=F, warning=F}
foreASD_result=as_tibble(read.csv("C:\\han\\Projects\\IntegratePPI\\forecASDPaper\\forecASD_table.csv"))

length(foreASD_result$symbol)

length(intersect(foreASD_result$symbol, jack_fu_185_asd_genes$gene_gencodeV33))
length(intersect(foreASD_result$symbol, jack_fu$gene_gencodeV33))
```


* foreASD outputs 17597 genes, of which 178 are in 185 risk genes in fu paper, and 15999 overlapping with all genes in fu paper.  



### gene symbol, ENSG, ENSP


```{r, message=F, warning=F}
#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")

#BiocManager::install("biomaRt")
library(biomaRt)


ensembl <- useMart("ensembl", dataset = "hsapiens_gene_ensembl", host = "https://grch37.ensembl.org")

# List of gene symbols
jack_fu_genes=jack_fu$gene_gencodeV33


# Define function to split queries
query_genes_in_batches <- function(genes, batch_size = 50) {
  results <- list()
  for (i in seq(1, length(genes), by = batch_size)) {
    batch <- genes[i:min(i + batch_size - 1, length(genes))]
    cat("Querying batch:", i, "to", min(i + batch_size - 1, length(genes)), "\n")
    
    batch_result <- tryCatch({
      getBM(attributes = c("hgnc_symbol", "ensembl_gene_id", "ensembl_peptide_id"),
            filters = "hgnc_symbol",
            values = batch,
            mart = ensembl)
    }, error = function(e) {
      cat("Error in batch query:", e$message, "\n")
      return(NULL)
    })
    
    results <- append(results, list(batch_result))
  }
  return(do.call(rbind, results))
}

# Run the batch query
jack_fu_gene_info <- query_genes_in_batches(jack_fu_genes, batch_size = 1000)
#write.csv(jack_fu_gene_info, file="C:\\Users\\hans\\OneDrive - Marquette University\\AI_for_Autism\\jack_fu_gene_info.csv")

bs_genes_info=query_genes_in_batches(bs_genes, batch_size = 1000)
#write.csv(bs_genes_info, file="C:\\Users\\hans\\OneDrive - Marquette University\\AI_for_Autism\\bs_genes_info.csv")



```




## train the model 


## results 


* [deep-motif score distribution](part01_score_dist.html)  

* [enrichment of top rank genes](part02_enrichment.html)

