---
title: "07/25 2025"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
date: "2025-07-25"
---



```{r, message=F, warning=F}
rm(list=ls())
set.seed(123)
library(tidyverse)
library(ggplot2)
library(DT)
library(VennDiagram)
library(rprojroot)
root <- rprojroot::find_rstudio_root_file()
```





```{r, message=F, warning=F}
#result_round1=as_tibble(read.csv(file.path(root, "../DiWangResults/ASD_prediction_results/ASD_RISK_PREDICTION_round1.csv"))) # old results 

result_round1=as_tibble(read.csv(file.path(root, "../DiWangResults/results updated_20250303/ASD_RISK_PREDICTION round 1.csv")))
result_round1=result_round1[result_round1['gene'] != 0,]  # remove rows with gene=0


SFARI_genes=as_tibble(read.csv(file.path(root, "../../../Dataset/SFARI_base/SFARI-Gene_genes_08-19-2024release_09-16-2024export.csv")))
positive_genes=SFARI_genes %>% filter (gene.score=="1") %>% dplyr::select(gene.symbol) %>% pull()
positive_genes=intersect(positive_genes, result_round1$gene)


score_cutoff=0.5
top_genes=result_round1 %>% filter(RISK_PROBABILITY_FUSION>score_cutoff) %>% dplyr::select(gene_gencodeV33) %>% pull

```



## clusterprofiler 


```{r, message=F, warning=F}

#if (!requireNamespace("BiocManager", quietly = TRUE))
#  install.packages("BiocManager")

#BiocManager::install(c("clusterProfiler", "org.Hs.eg.db", "enrichplot"))


# Load the libraries
library(clusterProfiler)
library(org.Hs.eg.db)  # For human gene annotations
library(enrichplot)    # For visualization



```




```{r, message=F, warning=F}
ego <- enrichGO(gene          = top_genes, 
               OrgDb         = org.Hs.eg.db, 
               keyType       = "SYMBOL", 
               ont           = "All",        # BP: Biological Process, CC: Cellular Component, MF: Molecular Function
               pAdjustMethod = "BH", 
               pvalueCutoff  = 0.05, 
               qvalueCutoff  = 0.01)

# View the result
#head(ego)

# Dotplot for GO enrichment
dotplot(ego, showCategory=nrow(ego), title="")+
  theme(axis.text.y = element_text(size = 6))  # Adjust the size as needed

# Extract significant terms while preserving the enrichResult object
#significant_ego <- ego
#significant_ego@result <- significant_ego@result[significant_ego@result$p.adjust < 0.05, ]

# Create the dot plot with significant terms only
#dotplot(significant_ego, showCategory = nrow(significant_ego), title = "GO Enrichment - Biological Process")
```



### Gene-Concept Network




```{r, message=F, warning=F}

# Convert gene symbols to Entrez IDs
gene_df <- bitr(top_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# Extract Entrez IDs
entrez_ids <- gene_df$ENTREZID
#print(entrez_ids)

ekegg <- enrichKEGG(gene         = entrez_ids, 
                   organism     = "hsa",      # Human
                   pAdjustMethod = "BH", 
                   pvalueCutoff  = 0.05)

# View the result
#head(ekegg)

# Barplot for KEGG enrichment
#barplot(ekegg, showCategory= nrow(ekegg), title="KEGG Pathway Enrichment")+
#  theme(axis.text.y = element_text(size = 5))


#Sort ekegg results by Count in descending order
ekegg_sorted <- ekegg
ekegg_sorted@result <- ekegg@result[order(ekegg@result$Count, decreasing = TRUE), ]

# Plot with sorted categories
barplot(ekegg_sorted,
        showCategory = nrow(ekegg_sorted),  # show all
        title = "KEGG Pathway Enrichment") +
  theme(axis.text.y = element_text(size = 5), 
        plot.title = element_text(hjust = 0.5))  # Center title

```


```{r, message=F, warning=F}
# Extract and filter
df <- as.data.frame(ekegg@result)
df_filtered <- df[df$p.adjust < 0.05, ]

# Sort by Fold Enrichment
df_filtered <- df_filtered[order(df_filtered$FoldEnrichment, decreasing = TRUE), ]

# Plot
ggplot(df_filtered, aes(x = reorder(Description, FoldEnrichment), 
                        y = FoldEnrichment,
                        fill = p.adjust)) +
  geom_col() +
  coord_flip() +
  scale_fill_gradient(low = "red", high = "blue", name = "Adj. p-value") +
  labs(title = "(A) Significant KEGG Pathways",
       x = "",
       y = "Fold Enrichment") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 25),
    axis.text.y = element_text(size = 10), 
    axis.text.x = element_text(size = 14), 
    legend.title = element_text(size = 14),      # increase legend title
    legend.text = element_text(size = 12),        # increase legend label text, 
     axis.title.x = element_text(size = 18)       # y-axis title (this line was added)
  )
length(unique(df_filtered$Description))

```


*  63 significant enriched deep-motifs pathways 



#### SFARI genes 

```{r, message=F, warning=F}

# Convert gene symbols to Entrez IDs
gene_df <- bitr(positive_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# Extract Entrez IDs
entrez_ids <- gene_df$ENTREZID
#print(entrez_ids)

sfari_ekegg <- enrichKEGG(gene         = entrez_ids, 
                   organism     = "hsa",      # Human
                   pAdjustMethod = "BH", 
                   pvalueCutoff  = 0.05)


# Extract and filter
sfari_df <- as.data.frame(sfari_ekegg@result)
sfari_df_filtered <- sfari_df[sfari_df$p.adjust < 0.05, ]

# Sort by Fold Enrichment
sfari_df_filtered <- sfari_df_filtered[order(sfari_df_filtered$FoldEnrichment, decreasing = TRUE), ]

# Plot
sfari_kegg_pathway=ggplot(sfari_df_filtered, aes(x = reorder(Description, FoldEnrichment), 
                        y = FoldEnrichment,
                        fill = p.adjust)) +
  geom_col() +
  coord_flip() +
  scale_fill_gradient(low = "red", high = "blue", name = "Adj. p-value") +
  labs(title = "SFARI",
       x = "",
       y = "Fold Enrichment") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 25),
    axis.text.y = element_text(size = 10), 
    axis.text.x = element_text(size = 14), 
    legend.title = element_text(size = 14),      # increase legend title
    legend.text = element_text(size = 12),        # increase legend label text, 
     axis.title.x = element_text(size = 18)       # y-axis title (this line was added)
  )
sfari_kegg_pathway
#kegg_pathway


length(unique(sfari_df_filtered$Description))
```


* 59 significant sfari enriched pathways  



#### compare SFARI vs MOTIFs

```{r, message=F, warning=F}
library(dplyr)
library(ggplot2)

# Add Source column
df_filtered$Source <- "Deep-MOTIFs"
sfari_df_filtered$Source <- "SFARI"

# Combine datasets
combined_df <- bind_rows(df_filtered, sfari_df_filtered)

# Replace NA counts with 0
combined_df$Count[is.na(combined_df$Count)] <- 0

# Step 1: Get Description order from SFARI (largest to smallest)
sfari_order <- combined_df %>%
  filter(Source == "SFARI") %>%
  arrange(desc(Count)) %>%
  pull(Description) %>%
  unique()

# Step 2: Add in Deep-MOTIFs-only descriptions to the bottom
all_desc <- unique(combined_df$Description)
deep_only_desc <- setdiff(all_desc, sfari_order)
final_levels <- c(sfari_order, deep_only_desc)

# Step 3: Apply new ordering to the Description factor
combined_df$Description <- factor(combined_df$Description, levels = final_levels)

# Step 4: Plot
ggplot(combined_df, aes(x = Description, y = Count, fill = Source)) +
  geom_col(position = position_dodge(width = 0.7)) +
  coord_flip() +
  labs(title = "(A) KEGG Pathway Comparison",
       x = "",
       y = "Gene Count") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 25),
    axis.text.y = element_text(size = 10), 
    axis.text.x = element_text(size = 14), 
    legend.title = element_text(size = 14),      
    legend.text = element_text(size = 12),        
    axis.title.x = element_text(size = 18)
  ) +
  guides(fill = guide_legend(title = NULL)) +
  scale_fill_manual(values = c("SFARI" = "steelblue", "Deep-MOTIFs" = "tomato"))



```

* 46 significant enriched pathways both in sfari and deep-motifs 



```{r, message=F, warning=F}
 ggplot(combined_df, aes(x = Description, 
                                                 y = -log(p.adjust),
                                                 fill = Source)) +
  geom_col(position = position_dodge(width = 0.7)) +
  coord_flip() +
  labs(title = "(A) KEGG Pathway Comparison",
       x = "",
       y = "-log(p value)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 25),
    axis.text.y = element_text(size = 10), 
    axis.text.x = element_text(size = 14), 
    legend.title = element_text(size = 14),      
    legend.text = element_text(size = 12),        
    axis.title.x = element_text(size = 18)
  ) +
  scale_fill_manual(values = c("SFARI" = "steelblue", "Deep-MOTIFs" = "tomato"))

```


```{r, message=F, warning=F}
#combined_df <- combined_df %>%
 # group_by(Source) %>%
#  mutate(Description = fct_reorder(Description, FoldEnrichment))

kegg_comparison_plot=ggplot(combined_df, aes(x = Description, y = FoldEnrichment, fill = p.adjust)) +
  geom_col() +
  facet_wrap(~Source) +
  scale_fill_gradient(low = "red", high = "blue", name = "Adj. p-value") +
  coord_flip() +
  labs(title = "(A)",
       x = "",
       y = "Fold enrichment") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 50),
    strip.text = element_text(size = 30, face = "bold"),  # facet subtitle size
    axis.text.y = element_text(size = 14), 
    axis.text.x = element_text(size = 35), 
    legend.title = element_text(size = 24),      
    legend.text = element_text(size = 22),        
    axis.title.x = element_text(size = 30),
    axis.title.y = element_blank()                # remove y-axis label
  )
kegg_comparison_plot
```









```{r, message=F, warning=F}
wide_df <- combined_df %>%
  dplyr::select(category, subcategory, Description, FoldEnrichment,p.adjust, Count, Source) %>%
  pivot_wider(
    names_from = Source,
    values_from = c(FoldEnrichment, p.adjust, Count),
    names_sep = "."
  )%>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))

wide_df%>% 
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```





```{r, message=F, warning=F}

# Convert gene symbols to Entrez IDs
gene_df <- bitr(top_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# Extract Entrez IDs
entrez_ids <- gene_df$ENTREZID
#print(entrez_ids)

ekegg <- enrichKEGG(gene         = entrez_ids, 
                   organism     = "hsa",      # Human
                   pAdjustMethod = "BH", 
                   pvalueCutoff  = 0.05)

# Network plot for KEGG pathways
cnetplot(ekegg, circular = TRUE, colorEdge = TRUE, cex_label_gene=0.5, cex_label_category=5)
length(unique(ekegg@gene))
```





```{r, message=F, warning=F}



# Convert Entrez IDs to Gene Symbols
gene_symbols <- bitr(entrez_ids, fromType = "ENTREZID", toType = "SYMBOL", OrgDb = org.Hs.eg.db)

# Merge the converted gene symbols back into the enrichment result
ekegg@result$geneID <- sapply(strsplit(ekegg@result$geneID, "/"), function(ids) {
  symbols <- gene_symbols$SYMBOL[match(ids, gene_symbols$ENTREZID)]
  paste(symbols, collapse = "/")
})
length(unique(ekegg@gene))


# Plot the cnetplot with gene symbols
cnetplot(ekegg, circular = TRUE, colorEdge = TRUE, cex_label_gene=0.4, cex_label_category=1)


```






* [reference-clusterProfiler 4.0](https://www.sciencedirect.com/science/article/pii/S2666675821000667)


```{r, message=F, warning=F}
# Your gene list (symbols or IDs from ekegg)
all_genes <- unique(unlist(strsplit(ekegg@result$geneID, split = "/")))

# Define new genes (example)
new_genes <- setdiff(all_genes, positive_genes)  # replace with your actual new genes

# Example: your gene groups vector (names are gene symbols or IDs)
gene_group <- ifelse(all_genes %in% new_genes, "new", "old")
names(gene_group) <- all_genes

# Create a vector of colors by gene group
group_colors <- c(new = "red", old = "grey50", category = "orange")


# Replace long description with line break version
ekegg@result$Description <- gsub(
  "ATP-dependent chromatin remodeling",
  "ATP-dependent\nchromatin remodeling",
  ekegg@result$Description
)


# Generate cnetplot and assign it to a variable
p <- cnetplot(ekegg,
              circular = TRUE,
              colorEdge = TRUE,
              cex_label_gene = 1.3,
              cex_label_category = 1)

# Assign 'node_group' based on 'name'
p$data <- p$data %>%
  mutate(node_group = case_when(
    name %in% new_genes ~ "new",
    name %in% positive_genes ~ "old",
    TRUE ~ "category"
  )) %>%
  # Map group to color column used by plot
  mutate(color = group_colors[node_group])

# Now force ggplot to use your colors for the nodes
net_plot=p + 
  # Override color scale to identity so colors in 'color' column are used directly
  scale_color_identity()+
  ggtitle("(B)")+  # Replace with your desired title
  theme(plot.title = element_text(hjust = 0.5, size=50), 
        legend.position = "right", 
        legend.title = element_blank(),                      # Remove legend title
    legend.text = element_text(size = 20)                # Bigger legend labels
    )
net_plot
```



#### top 5 pathways, 63 unique genes  

```{r, message=F, warning=F, eval=T}
# Extract top 5 pathways by p values 
ekegg_top5 <- ekegg
ekegg_top5@result <- ekegg@result[1:5, ]

# Convert to tibble to use dplyr functions safely
ekegg_df <- as_tibble(ekegg_top5@result)

# Extract gene–pathway relationships
gene_pathway_df <- ekegg_df %>%
  dplyr::select(Description, geneID) %>%
  tidyr::separate_rows(geneID, sep = "/") %>%
  dplyr::rename(pathway = Description, gene = geneID) %>%
  dplyr::distinct()


# Filter genes shown in the cnetplot
Genes_in_pathways <- unique(p$data$name[p$data$node_group %in% c("new", "old")])
gene_pathway_df <- gene_pathway_df %>%
  filter(gene %in% Genes_in_pathways)

# Add annotations
in_positive <- ifelse(gene_pathway_df$gene %in% positive_genes, "Yes", "NO")
in_deep_motifs <- ifelse(gene_pathway_df$gene %in% top_genes, "Yes", "NO")

names(in_positive) <- gene_pathway_df$gene
names(in_deep_motifs) <- gene_pathway_df$gene

annotation_df <- gene_pathway_df %>%
  mutate(
    IS_SFARI_gene = in_positive[gene],
    IS_MOTIFs_gene = in_deep_motifs[gene]
  )

annotation_df%>%
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

length(unique(annotation_df$gene))

```



















```{r, message=F, warning=F, eval=F}
library(gridExtra)
all <- grid.arrange(kegg_comparison_plot,  net_plot, nrow=1) 
ggsave(all,filename = paste0("C:\\Users\\hans\\OneDrive - Marquette University\\AI_for_Autism\\Fig\\KEGG-pathway-Deep-MOTIFs.pdf"),width = 29,height = 12,dpi = 600)
```



```{r, message=F, warning=F, eval=F}
all <- grid.arrange(ranking_change,  gene_set_enrichment,kegg_comparison_plot,  net_plot,  nrow=2) 
ggsave(all,filename = paste0("C:\\Users\\hans\\OneDrive - Marquette University\\AI_for_Autism\\Fig\\real_data_validation.pdf"),width = 20,height = 16,dpi = 900)
```




```{r, message=F, warning=F}
# show how many genes are displayed in the graph
top_n <- 5  # change this number to your desired showCategory
top_terms <- ekegg@result %>%
  arrange(p.adjust) %>%
  slice_head(n = top_n)

genes_shown <- unique(unlist(strsplit(top_terms$geneID, "/")))
length(genes_shown)  # Number of genes that will appear

novel_genes=intersect(genes_shown, new_genes)

length(novel_genes)

setdiff(novel_genes, SFARI_genes$gene.symbol)

```


* 63 genes were displayed in top 5 enriched pathways 

* 32 are novel gene by Deep-MOTIFs, 22 of which are not in any SFARI categories.  


#### top 20 enriched terms 


```{r, message=F, warning=F}
#library(dplyr)
#ekegg@result %>%
#  arrange(p.adjust) %>%
#  select(ID, Description, p.adjust) %>%
#  head(10)  # Top 10 enriched terms


#list top 10 enriched terms 
ekegg_result_sorted <- ekegg@result[order(ekegg@result$p.adjust), ]
head(ekegg_result_sorted[, c("ID", "Description", "p.adjust")], 20)
```





