---
title: "functional enrichment"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
date: "2025-03-18"
---


```{r, message=F, warning=F}
rm(list=ls())
set.seed(123)
library(tidyverse)
library(ggplot2)
library(DT)
library(VennDiagram)
library(rprojroot)
root <- rprojroot::find_rstudio_root_file()
```





```{r, message=F, warning=F}
#result_round1=as_tibble(read.csv(file.path(root, "../DiWangResults/ASD_prediction_results/ASD_RISK_PREDICTION_round1.csv"))) # old results 

result_round1=as_tibble(read.csv(file.path(root, "../DiWangResults/results updated_20250303/ASD_RISK_PREDICTION round 1.csv")))
result_round1=result_round1[result_round1['gene'] != 0,]  # remove rows with gene=0


SFARI_genes=as_tibble(read.csv(file.path(root, "../../../Dataset/SFARI_base/SFARI-Gene_genes_08-19-2024release_09-16-2024export.csv")))
positive_genes=SFARI_genes %>% filter (gene.score=="1") %>% dplyr::select(gene.symbol) %>% pull()
positive_genes=intersect(positive_genes, result_round1$gene)

```


## gprofiler 



### deep motif genes 


```{r, message=F, warning=F}
score_cutoff=0.5
top_genes=result_round1 %>% filter(RISK_PROBABILITY_FUSION>score_cutoff) %>% dplyr::select(gene_gencodeV33) %>% pull

#install.packages("gprofiler2")
library(gprofiler2)   ### https://biit.cs.ut.ee/gprofiler/page/r
gostres <- gost(query = top_genes,
organism = "hsapiens", significant = TRUE, correction_method = "bonferroni")

# output the enrichment 
gostres$result %>% 
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

dim(gostres$result)
```

* for 447 Deep-MOTIFs genes, there are 1722 significant enriched items. 



```{r, message=F, warning=F}

p <- gostplot(gostres, capped = F, interactive = F)
p
pp <- publish_gostplot(p, highlight_terms = c("HP:0000729", "HP:0025783", "HP:0000734", "HP:5200241", "GO:0007399"), 
                       width = NA, height = NA, filename = NULL )


```



### motif & SFARI genes 



```{r, message=F, warning=F}
#install.packages("gprofiler2")
library(gprofiler2)   ### https://biit.cs.ut.ee/gprofiler/page/r
gostres2 <- gost(query = list("Deep-MOTIFs genes"=top_genes, "SFARI genes"=positive_genes), 
organism = "hsapiens", significant = TRUE, correction_method = "bonferroni")
```

* in total 2960 significant enrichment items for 447 Deep-MOTIFs genes and 208 SFARI risk genes. 


```{r, message=F, warning=F}
gostplot(gostres2, capped = F, interactive = TRUE)
```




```{r, message=F, warning=F}
pvalue_threshold=0.05
#output_terms=gostres2$result$term_id[gostres2$result$p_value<1e-20]
#publish_gosttable(gostres2, 
#                         highlight_terms = output_terms,
#                        use_colors = TRUE, 
#                        show_columns = c("source", "term_name", "term_size"),
#                        filename = NULL)

highlight_terms=gostres2$result %>% filter(p_value<pvalue_threshold) %>% dplyr::select(term_id, term_name,  query, p_value)
length(highlight_terms)

highlight_terms_motif_term_id=highlight_terms %>% filter(query=="Deep-MOTIFs genes") %>% dplyr::select(term_id)%>% pull()
highlight_terms_sfari_term_id=highlight_terms %>% filter(query=="SFARI genes") %>% dplyr::select(term_id)%>% pull()
common_term_id=intersect(highlight_terms_motif_term_id, highlight_terms_sfari_term_id)
#highlight_terms %>% filter(term_id %in% common_term_id)
length(common_term_id)

```




```{r, message=F, warning=F}
ggplot(highlight_terms %>% filter(term_id %in% common_term_id), aes(x = term_id, y = -log10(p_value), fill = query)) +
  geom_bar(stat = "identity", position = position_dodge()) +
 # coord_flip() +  # Rotate x and y axes
  theme(axis.text.x = element_blank(),           # remove x-axis text
        axis.ticks.x = element_blank(),          # remove x-axis ticks 
        axis.text.y = element_text(size = 15), 
        axis.title.x = element_text(size = 15)   # Font size for y-axis label
        )+
  labs(title = "Comparison of top enriched items for Deep-MOTIFs genes and SFARI genes ", x = "", y = expression(paste(-log[10](p-value))))+
  theme(legend.title = element_blank()# Remove legend title
        )+
   theme(plot.title = element_text(hjust = 0.5, size=12))  #center the title

```




* [reference](https://cran.r-project.org/web/packages/gprofiler2/vignettes/gprofiler2.html#gene-list-functional-enrichment-analysis-with-gost)



```{r, message=F, warning=F}



highlight_terms_wide=highlight_terms%>%  # convert from long format to wide format 
  pivot_wider(
    names_from = query,
    values_from = p_value,
    names_prefix = "pvalue_"
  )
highlight_terms_wide=highlight_terms_wide %>% drop_na()  # remove enriched terms with NA  pvalue from either deep motifs or sfari 

dim(highlight_terms_wide)


highlight_terms_wide <- highlight_terms_wide %>%
  mutate(group = case_when(
    grepl("^GO:", term_id) ~ "GO",
    grepl("^HPA:", term_id) ~ "HPA",
    grepl("^HP:", term_id) ~ "HP",
    grepl("^TF:", term_id) ~ "TF",
    TRUE ~ "Others"
  ))

top10_labels <- highlight_terms_wide %>%
  mutate(min_p = pmin(`pvalue_Deep-MOTIFs genes`, `pvalue_SFARI genes`)) %>%
  arrange(min_p) %>%
  dplyr::slice(1:10)

library(ggrepel)

ggplot(highlight_terms_wide, aes(x = -log10(`pvalue_Deep-MOTIFs genes`), 
                                 y = -log10(`pvalue_SFARI genes`), 
                                 color = group)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "solid", color = "black", size=1) +
  scale_color_manual(values = c(
    "GO" = "goldenrod2", 
   # "HPA" = "purple4", 
   "HPA" = "black",
    "HP" = "990", 
    "TF" = "blue",
    "Others" = "grey70")) +
  geom_text_repel(
    data = top10_labels,
    aes(label = term_name),
    size = 3,
    max.overlaps = Inf
  ) +
  labs(
    title = "",
    x = "-log10(p value): Deep-MOTIFs",
    y = "-log10(p value): SFARI",
    color = "Group"
  ) +
  theme(legend.title = element_blank())+
  theme(plot.title = element_text(hjust = 0.5, size = 12))



```

* each point is one significant GO term 



```{r, message=F, warning=F}
highlight_terms_wide=highlight_terms%>%  # convert from long format to wide format 
  pivot_wider(
    names_from = query,
    values_from = p_value,
    names_prefix = "pvalue_"
  )
#highlight_terms_wide=highlight_terms_wide %>% drop_na()  # remove enriched terms with NA  pvalue from either deep motifs or sfari 

highlight_terms_wide %>%
  filter(is.na(`pvalue_Deep-MOTIFs genes`) & !is.na(`pvalue_SFARI genes`)) %>%
  arrange(`pvalue_SFARI genes`) %>%
  mutate(rank = row_number())%>% 
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))


highlight_terms_wide %>%
  filter(!is.na(`pvalue_Deep-MOTIFs genes`) & is.na(`pvalue_SFARI genes`)) %>%
  arrange(`pvalue_Deep-MOTIFs genes`) %>%
  mutate(rank = row_number())%>% 
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

highlight_terms_wide <- highlight_terms_wide %>%
  mutate(across(where(is.numeric), ~replace_na(., 1)))  # replace NA with 1 

dim(highlight_terms_wide)
length(unique(highlight_terms$term_id))



```



```{r, message=F, warning=F}
highlight_terms_wide <- highlight_terms_wide %>%
  mutate(group = case_when(
    grepl("^GO:", term_id) ~ "GO",
    grepl("^HPA:", term_id) ~ "HPA",
    grepl("^HP:", term_id) ~ "HP",
    grepl("^TF:", term_id) ~ "TF",
    TRUE ~ "Others"
  ))

top10_labels <- highlight_terms_wide %>%
  mutate(min_p = pmin(`pvalue_Deep-MOTIFs genes`, `pvalue_SFARI genes`)) %>%
  arrange(min_p) %>%
  dplyr::slice(1:10)

library(ggrepel)

ggplot(highlight_terms_wide, aes(x = -log10(`pvalue_Deep-MOTIFs genes`), 
                                 y = -log10(`pvalue_SFARI genes`), 
                                 color = group)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "solid", color = "black", size=1) +
  scale_color_manual(values = c(
    "GO" = "goldenrod2", 
   # "HPA" = "purple4", 
   "HPA" = "black",
    "HP" = "990", 
    "TF" = "blue",
    "Others" = "grey70")) +
  geom_text_repel(
    data = top10_labels,
    aes(label = term_name),
    size = 3,
    max.overlaps = Inf
  ) +
  labs(
    title = "",
    x = "-log10(p value): Deep-MOTIFs",
    y = "-log10(p value): SFARI",
    color = "Group"
  ) +
  theme(legend.title = element_blank())+
  theme(plot.title = element_text(hjust = 0.5, size = 12))
```





## clusterprofiler 


```{r, message=F, warning=F}

#if (!requireNamespace("BiocManager", quietly = TRUE))
#  install.packages("BiocManager")

#BiocManager::install(c("clusterProfiler", "org.Hs.eg.db", "enrichplot"))


# Load the libraries
library(clusterProfiler)
library(org.Hs.eg.db)  # For human gene annotations
library(enrichplot)    # For visualization



```




```{r, message=F, warning=F}
ego <- enrichGO(gene          = top_genes, 
               OrgDb         = org.Hs.eg.db, 
               keyType       = "SYMBOL", 
               ont           = "All",        # BP: Biological Process, CC: Cellular Component, MF: Molecular Function
               pAdjustMethod = "BH", 
               pvalueCutoff  = 0.05, 
               qvalueCutoff  = 0.01)

# View the result
#head(ego)

# Dotplot for GO enrichment
dotplot(ego, showCategory=nrow(ego), title="")+
  theme(axis.text.y = element_text(size = 6))  # Adjust the size as needed

# Extract significant terms while preserving the enrichResult object
#significant_ego <- ego
#significant_ego@result <- significant_ego@result[significant_ego@result$p.adjust < 0.05, ]

# Create the dot plot with significant terms only
#dotplot(significant_ego, showCategory = nrow(significant_ego), title = "GO Enrichment - Biological Process")
```




```{r, message=F, warning=F, eval=F}
gsego <- gseGO(gene          = sort(top_genes, decreasing = T),  
               OrgDb         = org.Hs.eg.db,
               keyType       = "SYMBOL", 
               ont           = "All",        # BP: Biological Process, CC: Cellular Component, MF: Molecular Function
               pAdjustMethod = "BH", 
               pvalueCutoff  = 0.05, 
               qvalueCutoff  = 0.01)

# View the result
#head(ego)

# Dotplot for GO enrichment
dotplot(ego, showCategory=nrow(ego), title="")+
  theme(axis.text.y = element_text(size = 6))  # Adjust the size as needed

# Extract significant terms while preserving the enrichResult object
#significant_ego <- ego
#significant_ego@result <- significant_ego@result[significant_ego@result$p.adjust < 0.05, ]

# Create the dot plot with significant terms only
#dotplot(significant_ego, showCategory = nrow(significant_ego), title = "GO Enrichment - Biological Process")
```





```{r, message=F, warning=F, eval=F}

# Compute pairwise term similarity
ego_sim <- pairwise_termsim(ego)

# Visualize the results using the enrichment map plot
emapplot(ego_sim, label="ID")+
 # theme(legend.position = "none")
  theme(legend.text = element_text(size = 10))


emap <- emapplot(ego_sim, showCategory = 30, label =F)  # or any number of categories
emap + theme(legend.text = element_text(size = 10))


```


### Gene-Concept Network




```{r, message=F, warning=F}

# Convert gene symbols to Entrez IDs
gene_df <- bitr(top_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# Extract Entrez IDs
entrez_ids <- gene_df$ENTREZID
#print(entrez_ids)

ekegg <- enrichKEGG(gene         = entrez_ids, 
                   organism     = "hsa",      # Human
                   pAdjustMethod = "BH", 
                   pvalueCutoff  = 0.05)

# View the result
#head(ekegg)

# Barplot for KEGG enrichment
barplot(ekegg, showCategory= nrow(ekegg), title="KEGG Pathway Enrichment")+
  theme(axis.text.y = element_text(size = 5))


#Sort ekegg results by Count in descending order
ekegg_sorted <- ekegg
ekegg_sorted@result <- ekegg@result[order(ekegg@result$Count, decreasing = TRUE), ]

# Plot with sorted categories
barplot(ekegg_sorted,
        showCategory = nrow(ekegg_sorted),  # show all
        title = "KEGG Pathway Enrichment") +
  theme(axis.text.y = element_text(size = 5), 
        plot.title = element_text(hjust = 0.5))  # Center title

```


```{r, message=F, warning=F}
# Extract and filter
df <- as.data.frame(ekegg@result)
df_filtered <- df[df$p.adjust < 0.05, ]

# Sort by Fold Enrichment
df_filtered <- df_filtered[order(df_filtered$FoldEnrichment, decreasing = TRUE), ]

# Plot
ggplot(df_filtered, aes(x = reorder(Description, FoldEnrichment), 
                        y = FoldEnrichment,
                        fill = p.adjust)) +
  geom_col() +
  coord_flip() +
  scale_fill_gradient(low = "red", high = "blue", name = "Adj. p-value") +
  labs(title = "KEGG Pathway Enrichment (p.adjust < 0.05)",
       x = "",
       y = "Fold Enrichment") +
  #theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.y = element_text(size = 6)
  )
```






```{r, message=F, warning=F}



# Network plot for KEGG pathways
cnetplot(ekegg, circular = TRUE, colorEdge = TRUE, cex_label_gene=0.5, cex_label_category=5)
length(unique(ekegg@gene))
```





```{r, message=F, warning=F}



# Convert Entrez IDs to Gene Symbols
gene_symbols <- bitr(entrez_ids, fromType = "ENTREZID", toType = "SYMBOL", OrgDb = org.Hs.eg.db)

# Merge the converted gene symbols back into the enrichment result
ekegg@result$geneID <- sapply(strsplit(ekegg@result$geneID, "/"), function(ids) {
  symbols <- gene_symbols$SYMBOL[match(ids, gene_symbols$ENTREZID)]
  paste(symbols, collapse = "/")
})
length(unique(ekegg@gene))


# Plot the cnetplot with gene symbols
cnetplot(ekegg, circular = TRUE, colorEdge = TRUE, cex_label_gene=0.4, cex_label_category=1)


```






* [reference-clusterProfiler 4.0](https://www.sciencedirect.com/science/article/pii/S2666675821000667)


```{r, message=F, warning=F}
# Your gene list (symbols or IDs from ekegg)
all_genes <- unique(unlist(strsplit(ekegg@result$geneID, split = "/")))

# Define new genes (example)
new_genes <- setdiff(all_genes, positive_genes)  # replace with your actual new genes

# Example: your gene groups vector (names are gene symbols or IDs)
gene_group <- ifelse(all_genes %in% new_genes, "new", "old")
names(gene_group) <- all_genes

# Create a vector of colors by gene group
group_colors <- c(new = "red", old = "grey50", category = "orange")


# Generate cnetplot and assign it to a variable
p <- cnetplot(ekegg,
              circular = TRUE,
              colorEdge = TRUE,
              cex_label_gene = 0.4,
              cex_label_category = 1)

# Assign 'node_group' based on 'name'
p$data <- p$data %>%
  mutate(node_group = case_when(
    name %in% new_genes ~ "new",
    name %in% positive_genes ~ "old",
    TRUE ~ "category"
  )) %>%
  # Map group to color column used by plot
  mutate(color = group_colors[node_group])

# Now force ggplot to use your colors for the nodes
p + 
  # Override color scale to identity so colors in 'color' column are used directly
  scale_color_identity()

```





```{r, message=F, warning=F}
# show how many genes are displayed in the graph
top_n <- 5  # change this number to your desired showCategory
top_terms <- ekegg@result %>%
  arrange(p.adjust) %>%
  slice_head(n = top_n)

genes_shown <- unique(unlist(strsplit(top_terms$geneID, "/")))
length(genes_shown)  # Number of genes that will appear

novel_genes=intersect(genes_shown, new_genes)

length(novel_genes)

setdiff(novel_genes, SFARI_genes$gene.symbol)

```


* 63 genes were displayed in top 5 enriched pathways 

* 32 are novel gene by Deep-MOTIFs, 22 of which are not in any SFARI categories.  





```{r, message=F, warning=F}
#library(dplyr)
#ekegg@result %>%
#  arrange(p.adjust) %>%
#  select(ID, Description, p.adjust) %>%
#  head(10)  # Top 10 enriched terms


#list top 10 enriched terms 
ekegg_result_sorted <- ekegg@result[order(ekegg@result$p.adjust), ]
head(ekegg_result_sorted[, c("ID", "Description", "p.adjust")], 20)
```




```{r, message=F, warning=F, eval=F}
pdf("cnetplot_all56terms.pdf", width = 16, height = 12)
cnetplot(ekegg, showCategory = 56, circular = TRUE, colorEdge = TRUE)
dev.off()
```



## novel pathways in MOTIFs


```{r, message=F, warning=F}
pvalue_threshold=0.05
gostres2_all <- gost(query = list("Deep-MOTIFs genes"=top_genes, "SFARI genes"=positive_genes), 
organism = "hsapiens", significant = F, correction_method = "bonferroni")

```



```{r, message=F, warning=F}
highlight_terms_motif_term_id=gostres2_all$result %>% filter(query=="Deep-MOTIFs genes")%>% filter(p_value<pvalue_threshold)  %>% dplyr::select(term_id)%>% pull()
highlight_terms_sfari_term_id=gostres2_all$result %>% filter(query=="SFARI genes")%>% filter(p_value>pvalue_threshold)%>% dplyr::select(term_id)%>% pull()
common_term_id=intersect(highlight_terms_motif_term_id, highlight_terms_sfari_term_id)


highlight_terms=gostres2_all$result %>% filter (term_id %in% common_term_id) %>% dplyr::select(term_id, term_name, query, p_value)
```





```{r, message=F, warning=F}
highlight_terms_wide=highlight_terms%>%  # convert from long format to wide format 
  pivot_wider(
    names_from = query,
    values_from = p_value,
    names_prefix = "pvalue_"
  )
highlight_terms_wide=highlight_terms_wide %>% drop_na()
dim(highlight_terms_wide)

highlight_terms_wide%>% 
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

```



```{r, message=F, warning=F}
highlight_terms_wide <- highlight_terms_wide %>%
  mutate(group = case_when(
    grepl("^GO:", term_id) ~ "GO",
    grepl("^HPA:", term_id) ~ "HPA",
    grepl("^HP:", term_id) ~ "HP",
    grepl("^TF:", term_id) ~ "TF",
    TRUE ~ "Others"
  ))

top10_labels <- highlight_terms_wide %>%
  mutate(min_p = pmin(`pvalue_Deep-MOTIFs genes`, `pvalue_SFARI genes`)) %>%
  arrange(min_p) %>%
  dplyr::slice(1:10)

library(ggrepel)

ggplot(highlight_terms_wide, aes(x = -log10(`pvalue_Deep-MOTIFs genes`), 
                                 y = -log10(`pvalue_SFARI genes`), 
                                 color = group)) +
  geom_point(alpha = 0.6, size = 2) +
  #geom_abline(slope = 1, intercept = 0, linetype = "solid", color = "black", size=1) +
  scale_color_manual(values = c(
    "GO" = "goldenrod2", 
    "HPA" = "purple4", 
    "HP" = "990", 
    "TF" = "blue",
    "Others" = "grey70")) +
  geom_text_repel(
    data = top10_labels,
    aes(label = term_name),
    size = 3,
    max.overlaps = Inf
  ) +
  labs(
    title = "",
    x = "-log10(p value): Deep-MOTIFs",
    y = "-log10(p value): SFARI",
    color = "Group"
  ) +
  theme(legend.title = element_blank())+
  theme(plot.title = element_text(hjust = 0.5, size = 12))

```



## missed pathways in MOTIFs


```{r, message=F, warning=F}
pvalue_threshold=0.05
gostres2_all <- gost(query = list("Deep-MOTIFs genes"=top_genes, "SFARI genes"=positive_genes), 
organism = "hsapiens", significant = F, correction_method = "bonferroni")

```



```{r, message=F, warning=F}
highlight_terms_motif_term_id=gostres2_all$result %>% filter(query=="Deep-MOTIFs genes")%>% filter(p_value>pvalue_threshold)  %>% dplyr::select(term_id)%>% pull()
highlight_terms_sfari_term_id=gostres2_all$result %>% filter(query=="SFARI genes")%>% filter(p_value<pvalue_threshold)%>% dplyr::select(term_id)%>% pull()
common_term_id=intersect(highlight_terms_motif_term_id, highlight_terms_sfari_term_id)


highlight_terms=gostres2_all$result %>% filter (term_id %in% common_term_id) %>% dplyr::select(term_id, term_name, query, p_value)
```





```{r, message=F, warning=F}
highlight_terms_wide=highlight_terms%>%  # convert from long format to wide format 
  pivot_wider(
    names_from = query,
    values_from = p_value,
    names_prefix = "pvalue_"
  )
highlight_terms_wide=highlight_terms_wide %>% drop_na()
dim(highlight_terms_wide)

highlight_terms_wide%>% 
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

```



```{r, message=F, warning=F}
highlight_terms_wide <- highlight_terms_wide %>%
  mutate(group = case_when(
    grepl("^GO:", term_id) ~ "GO",
    grepl("^HPA:", term_id) ~ "HPA",
    grepl("^HP:", term_id) ~ "HP",
    grepl("^TF:", term_id) ~ "TF",
    TRUE ~ "Others"
  ))

top10_labels <- highlight_terms_wide %>%
  mutate(min_p = pmin(`pvalue_Deep-MOTIFs genes`, `pvalue_SFARI genes`)) %>%
  arrange(min_p) %>%
  dplyr::slice(1:10)

library(ggrepel)

ggplot(highlight_terms_wide, aes(x = -log10(`pvalue_Deep-MOTIFs genes`), 
                                 y = -log10(`pvalue_SFARI genes`), 
                                 color = group)) +
  geom_point(alpha = 0.6, size = 2) +
  #geom_abline(slope = 1, intercept = 0, linetype = "solid", color = "black", size=1) +
  scale_color_manual(values = c(
    "GO" = "goldenrod2", 
    "HPA" = "purple4", 
    "HP" = "990", 
    "TF" = "blue",
    "Others" = "grey70")) +
  geom_text_repel(
    data = top10_labels,
    aes(label = term_name),
    size = 3,
    max.overlaps = Inf
  ) +
  labs(
    title = "",
    x = "-log10(p value): Deep-MOTIFs",
    y = "-log10(p value): SFARI",
    color = "Group"
  ) +
  theme(legend.title = element_blank())+
  theme(plot.title = element_text(hjust = 0.5, size = 12))

```